\section{Block Codes}

\subsection{Exercise 1}

\begin{formulationBox}
	Consider the code of words of the form $(m_1, m_2, m_3, m_4, r_1, r_2, r_3, r_4)$, where $(m_1, m_2, m_3, m_4)$ is the binary message we want to transmit and the redundant bits $(r_1, r_2, r_3, r_4)$ are defined by the following conditions. The sum of the number of ones in rows and columns 1 to 2 should be even.
	
	\begin{center}
		\begin{tabular}{|c|c|c|}
			\hline
			$m_1$ & $m_2$ & $r_1$\\
			\hline
			$m_3$ & $m_4$ & $r_2$\\
			\hline
			$r_3$ & $r_4$ \\
			\cline{1-2}
		\end{tabular},
	\end{center}
	
	Show that this code corrects one error.
\end{formulationBox}

If an error occurs in one of the of the message's bits, then the parity check in that row and that column will fail, so the error is located at those coordinates, and by flipping the bit the error is corrected.

If an error occurs in $r_3$ or $r_4$, then the parity check in its column will fail, but row parity checks will be correct, hence by elimination the error is located at the redundant bit's row, and by flipping the bit the error is corrected.

The reasoning for $r_1$ and $r_2$ is virtually the same.

$\qed$

\subsection{Exercise 2}

\begin{formulationBox}
	Let $C$ be the code of all $n \times n$ binary arrays that have an even number of ones in every row and column. What is the number of codewords of $C$? What is the minimum distance of $C$? Show that every pattern of three erasures can be corrected.
\end{formulationBox}

Let's first prove there exists exactly one codeword for every $(n-1)\times(n-1)$ binary array.

Take any binary array
\[A = \begin{pmatrix}
	a_{1,1} & a_{1,2} & \cdots & a_{1,n-1} \\
	a_{2,1} & a_{2,2} & \cdots & a_{2,n-1} \\
	\cdots & \cdots & \ddots & \vdots \\
	a_{n-1,1} & a_{n-1,2} & \cdots & a_{n-1,n-1} \\
\end{pmatrix}.\]
Then, we can build up an array
\[A' = \begin{pmatrix}
	a_{1,1} & a_{1,2} & \cdots & a_{1,n-1} & a_{1,n} \\
	a_{2,1} & a_{2,2} & \cdots & a_{2,n-1} & a_{2,n}\\
	\cdots & \cdots & \ddots & \vdots & \vdots \\
	a_{n-1,1} & a_{n-1,2} & \cdots & a_{n-1,n-1} & a_{n-1,n} \\
	a_{n,1} & a_{n,2} & \cdots & a_{n,n-1} & a_{n,n} \\
\end{pmatrix}\]
such that $A' \in C$, by setting the elements in the $n^\textrm{th}$ row and column to be equal to 1 when the number of ones in rest of elements is odd, and 0 otherwise. Likewise, if a bit in the $n^\textrm{th}$ row or column were flipped, then a bit in another row or column would need to be flipped as well, in order for the parity to be preserved.

This shows that the number of codewords of $C$ is exactly the number of possible $(n-1)\times(n-1)$ binary arrays; that is, $2^{(n-1)(n-1)}$.

The minimum distance of $C$ is equal to $2$; you can flip two elements in the same row or column preserving the parity, but that's not possible with a single flip.

Finally, an erasure can be corrected if the remaining elements in its row or its column have all be received. Let an erasure be at row $i$ and column $j$. Then, the worst case scenario would be having one other erasure in row $i$ and one other in column $j$, which would at first not let us correct the erasure at $i, j$. But we could then take one of the other erasures, say the one at $i, j'$ without loss of generality, and column $j'$ must have all the rest of elements set. This allows the erasure at $i, j'$ to be corrected first, and we would then be able to correct the erasure at $i, j$ and $i', j$.

